Global search algorithms for one-dimensional multiextremal optimization problems with nonlinear constraints

### 2017-09-19 2017/ParOptBook/Chapter4.tex
parOptBook: fix some infrastructure things in Chapter4

\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom
\usepackage{mathtools}
\mathtoolsset{showonlyrefs}

@@

\example{
\label{ex4:problem}
Let us consider a problem of type \eqref{eq4:problem} where $x\in [0.6;2.2],m=2$,
\begin{gather}
  \varphi(x)=\cos(18x-3)\sin(10x-7)+1.5, \\
  g_1(x)=\exp(-x/2)\sin(6x-1.5), \\
  g_2(x)=|x|\sin(2\pi x-0.5).
\end{gather}
he exact solution of the problem is $x^*=2.0795,\varphi(x^*)=0.565$. The graphs of the functions $g_1(x),g_2(x)$,and $\varphi(x)$ are presented in Fig. \ref{fig:4_1}.
\begin{figure}[h]
  \label{fig:4_1}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_1.png}
  \caption{One-dimensional global optimization problem}
\end{figure}
}

@@

\emph{Rule 6.} For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$:
\[
R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
\]
\begin{equation}
\label{eq4:characteristic}
R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
\end{equation}
\[
R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
\]
\[
\Delta_i=x_i - x_{i-1}
\]
The values $r_\nu>1, 1\le\nu\le m+1$ are the parameters of the algorithm. An appropriate choice of the values $r_\nu$ allows using the products $r_\nu\mu_\nu$ as the estimates of Lipschitz constants $L_\nu, 1\le\nu\le m+1$.
\emph{Rule 7.} Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
\begin{equation}
\label{eq4:MaxR}
R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\end{equation}
\emph{Rule 8.} Make the next trial at the midpoint of the interval
$(x_{t-1},x_t)$ if the indices of the points $x_{t-1}$ and $x_t$  are not the same, i.e.
\[
x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
\]
Otherwise, make the trial at the point
\begin{equation}
\label{eq4:next_point}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{z_t-z_{t-1}}{2r_\nu\mu},\; \nu=\nu(x_{t-1})=\nu(x_t).
\end{equation}

@@

\begin{table}

\begin{center}
  \label{tab4:exp_results}
  \begin{tabular}{| l | c | c | c| }
    \hline
    & $k_1$ & $k_2$ & $k_3$ \\ \hline
    Scanning a uniform grid & 160000 & 90280 & 56476 \\ \hline
    Penalty Function Method &  375 & 375 & 375 \\ \hline
    Index method & 63 & 49 & 35 \\ \hline
  \end{tabular}
\end{center}
\end{table}

>>>>

\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom

%custom package
\usepackage{mathtools}

@@

\example{
\label{ex4:problem}
Let us consider a problem of type \eqref{eq4:problem} where $x\in [0.6;2.2],m=2$,
\begin{gather*}
  \varphi(x)=\cos(18x-3)\sin(10x-7)+1.5, \\
  g_1(x)=\exp(-x/2)\sin(6x-1.5), \\
  g_2(x)=|x|\sin(2\pi x-0.5).
\end{gather*}
he exact solution of the problem is $x^*=2.0795,\varphi(x^*)=0.565$. The graphs of the functions $g_1(x),g_2(x)$,and $\varphi(x)$ are presented in Fig. \ref{fig:4_1}.
\begin{figure}[h]
  \label{fig:4_1}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_1.png}
  \caption{One-dimensional global optimization problem}
\end{figure}
}

@@

\emph{Rule 6.} For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$:
\begin{gather}
  \label{eq4:characteristic}
  R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}), \nonumber \\
  R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),\\
  R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i), \nonumber \\
  \Delta_i=x_i - x_{i-1} \nonumber
\end{gather}
The values $r_\nu>1, 1\le\nu\le m+1$ are the parameters of the algorithm. An appropriate choice of the values $r_\nu$ allows using the products $r_\nu\mu_\nu$ as the estimates of Lipschitz constants $L_\nu, 1\le\nu\le m+1$.

\emph{Rule 7.} Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
\begin{equation}
\label{eq4:MaxR}
R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\end{equation}

\emph{Rule 8.} Make the next trial at the midpoint of the interval
$(x_{t-1},x_t)$ if the indices of the points $x_{t-1}$ and $x_t$  are not the same, i.e.
\[
x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
\]
Otherwise, make the trial at the point
\begin{equation}
\label{eq4:next_point}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{z_t-z_{t-1}}{2r_\nu\mu},\; \nu=\nu(x_{t-1})=\nu(x_t).
\end{equation}

@@

\begin{table}
  \caption{}
\begin{center}
  \label{tab4:exp_results}
  \begin{tabular}{| l | c | c | c| }
    \hline
    & $k_1$ & $k_2$ & $k_3$ \\ \hline
    Scanning a uniform grid & 160000 & 90280 & 56476 \\ \hline
    Penalty Function Method &  375 & 375 & 375 \\ \hline
    Index method & 63 & 49 & 35 \\ \hline
  \end{tabular}
\end{center}
\end{table}
